SMT-Boosted Security Types for Low-Level MPC

### 2025-11-24 toplas26/intro.tex
attempt at intro

Our work is based on the $\metaprot$/$\minicat$ framework developed in
previous work \cite{skalka-near-ppdp24}. The prior work developed the 
language and semantics, plus two mechanisms for verifying security: an
automated approach for protocols in $\mathbb{F}_2$ (the binary field),
based on enumerating all possible executions of the protocol, that scales 
only to very small protocols; and a manual approach based on a program logic.
%
In Section \ref{section-lang} we recall the $\minicat$ language model and key definitions. In
Section \ref{section-model} we recall formulations of program distributions,
and key probabilistic hyperproperties of MPC in $\minicat$.

Together, our contributions \emph{automate} the verification of security
properties in $\metaprot$/$\minicat$. In this work, we build on~\cite{skalka-near-ppdp24}
by developing automatic verification approaches that work for arbitrary
finite fields and scale to large protocols. To accomplish this, we develop new
type systems for compositional verification, and use SMT to automate the checks
that were accomplished by enumeration in prior work. Specifically, we make the
following contributions:
%
\begin{enumerate}[i.]
\item In Section \ref{section-smt}, an interpretation of protocols in
  $\minicat$ as SMT constraints (Theorem \ref{theorem-toeq}) and
  methodology for verifying correctness properties.
\item In Section \ref{section-cpj}, a confidentiality type system for
  $\minicat$ protocols with a soundness property guaranteeing that
  information is released only through explicit declassifications
  (Theorem \ref{theorem-cpj}).
\item In Section \ref{section-ipj}, an integrity type system for
  $\minicat$ protocols with a soundness property guaranteeing
  robustness in malicious settings (Theorem \ref{theorem-ipj}).
\item In Section \ref{section-metalang}, a dependent Hoare type system and
  algorithm for the $\metaprot$ language that is sound for the
  confidentiality and integrity type systems (Theorem
  \ref{theorem-mtj}), with independently verified, compositional $\Pi$
  types for functions.
\end{enumerate}
In all cases, our type systems are boosted by verification mechanisms
provided by Satisfiability Modulo Finite Fields \cite{SMFF}. In
Section \ref{section-examples} we develop and discuss extended example
applications of our type analyses to real protocols including
the Goldreich-Micali-Wigderson (GMW), Bendlin-Damgard-Orland-Zakarias
(BDOZ), and Yao's Garbled Circuits (YGC) protocols.

>>>>

Our work is based on the $\metaprot$/$\minicat$ framework developed in
previous work \cite{skalka-near-ppdp24, skalka-near-esop25}. The prior work developed the 
language and semantics, plus two mechanisms for verifying security: an
automated approach for protocols in $\mathbb{F}_2$ (the binary field),
based on enumerating all possible executions of the protocol, that scales 
only to very small protocols; and a manual approach based on a program logic.
%
In Section \ref{section-lang} we recall the $\minicat$ language model and key definitions. In
Section \ref{section-model} we recall formulations of program distributions,
and key probabilistic hyperproperties of MPC in $\minicat$.

Together, our contributions \emph{automate} the \emph{compositional} verification of security
properties in $\metaprot$/$\minicat$. In this work, we build on~\cite{skalka-near-ppdp24}
by developing automatic verification approaches that work for arbitrary
finite fields and scale to large protocols. To accomplish this, we develop new
Hoare logics and type systems for compositional verification, and use SMT to automate the checks
that were accomplished by enumeration in prior work. Specifically, we make the
following contributions:
%
% \begin{enumerate}[i.]
% \item In Section \ref{section-smt}, an interpretation of protocols in
%   $\minicat$ as SMT constraints (Theorem \ref{theorem-toeq}) and
%   methodology for verifying correctness properties.
% \item In Section \ref{section-cpj}, a confidentiality type system for
%   $\minicat$ protocols with a soundness property guaranteeing that
%   information is released only through explicit declassifications
%   (Theorem \ref{theorem-cpj}).
% \item In Section \ref{section-ipj}, an integrity type system for
%   $\minicat$ protocols with a soundness property guaranteeing
%   robustness in malicious settings (Theorem \ref{theorem-ipj}).
% \item In Section \ref{section-metalang}, a dependent Hoare type system and
%   algorithm for the $\metaprot$ language that is sound for the
%   confidentiality and integrity type systems (Theorem
%   \ref{theorem-mtj}), with independently verified, compositional $\Pi$
%   types for functions.
% \end{enumerate}
\begin{enumerate}[i.]
\item In Section~\ref{section-smt}, a Hoare logic for Overture protocols, for the compositional verification of both semi-honest and malicious security.
\item In Section~\ref{section-sty}, a novel share type system, enabling security verification for protocols that use secret sharing as a building block; our approach works for any secret sharing scheme and enforces noninterference modulo output.
\item In Section~\ref{section-ipj}, an integrity type system for extending semi-honest security to malicious security.
\item In Section~\ref{section-metalang}, a Hoare logic for Prelude programs with an automated implementation, for the compositional verification of high-level protocol specifications.
\item In Section~\ref{sec:metapr-type-check}, a compositional type-based verification approach for Prelude programs that combines share types with the automated Hoare logic of Section 7.
\end{enumerate}
%
A prior conference version of this work~\cite{skalka-near-esop25}
proposed a whole-program analysis for Prelude programs that ensures
confidentiality, and combined it with an integrity type system to
prove security of Prelude programs. This extended version adds the
compositional Hoare logics presented in Sections 4 and 7, adds the
share type system presented in Section 5, and fully develops the
compositional type-based verification approach in Section 8. In
addition, we develop an implementation capable of automating the
verification of examples given in the prior work, and our extensions
in this work additionally enable the type-based verification of
noninterference modulo output for secret sharing protocols that
leverage Beaver triples (see Section~\ref{section-examples}).

In all cases, our type systems are boosted by verification mechanisms
provided by Satisfiability Modulo Finite Fields \cite{SMFF}. In
Section \ref{section-examples} we develop and discuss extended example
applications of our type analyses to real protocols including
the Goldreich-Micali-Wigderson (GMW), Bendlin-Damgard-Orland-Zakarias
(BDOZ), and Yao's Garbled Circuits (YGC) protocols.
